sWGS validation report

### 2019-09-10 Raw Data Comparison.Rmd
...

---
title: "Raw Data Comparison"
author: "Sarah Killcoyne"
date: "23/08/2019"
output: 
  html_document: 
    fig_height: 4
    fig_width: 6
---

@@

val.file = '~/Data/BarrettsProgressionRisk/QDNAseq/validation/sWGS_validation_batches.xlsx'
sheets = readxl::excel_sheets(val.file)[8:13]
all.val = do.call(bind_rows, lapply(sheets, function(s) {
  readxl::read_xlsx(val.file, s) %>% dplyr::select(`Hospital Research ID`, matches('Status'), `Block ID`,`Sample Type`, `SLX-ID`, `Index Sequence`, Cohort, Batch, RA, matches('Collection')) %>% dplyr::filter(!is.na(`SLX-ID`)) %>% mutate_at(vars(`SLX-ID`, `Block ID`), list(as.character))# %>% fncols('Collection', 'Biopsy')
})) %>% mutate(Samplename = paste(`SLX-ID`,`Index Sequence`, sep='.'))

@@

```{r, fig.width=12, fig.height=12}

@@

But what if I, as a quick test, normalize the validation cohort with itself? This looks better here, but when I use this for predictions it doesn't significantly change the results.

@@

@@

preds = left_join(all.val, preds, by=c('Samplename'='Sample', 'Hospital Research ID', 'SLX-ID', 'Index Sequence', 'Block ID'))

@@

How does it change if I use a global variance, or a rolling-bin variance (to preserve some local info) to adjust?

```{r}
# global variance
all.tiles = bind_rows(training.tiles, val.tiles)

means = apply(all.tiles[,-1], 2, mean)
sds = apply(all.tiles[,-1], 2, sd)
global.sd = sd(as.matrix(all.tiles[,-1]))

wt.sds = sapply(sds, function(x) weighted.mean(x,global.sd,w=c(0.2)))

plot(wt.sds, sds)

@@

@@

@@

@@

From training data
```{r}
plot(tx.normal)
```

From validation data
```{r}
plot(vx.normal)
```

>>>>

---
title: "Raw Data Comparison"
author: "Sarah Killcoyne"
date: "23/08/2019"
output: 
  html_document: 
    fig_height: 4
    fig_width: 6
    toc: yes
---

@@

val.file = '~/Data/BarrettsProgressionRisk/QDNAseq/validation/sWGS_validation_batches.xlsx'
sheets = readxl::excel_sheets(val.file)[8:13]

pastefun<-function(x) {
  if ( !grepl('SLX-', x) ) x = paste0('SLX-',x)
  return(x)
}

all.val = do.call(bind_rows, lapply(sheets, function(s) {
  readxl::read_xlsx(val.file, s) %>% dplyr::select(`Hospital Research ID`, matches('Status'), `Block ID`,`Sample Type`, `SLX-ID`, `Index Sequence`, Cohort, Batch, RA, matches('Collection')) %>% 
    dplyr::filter(!is.na(`SLX-ID`)) %>% mutate_at(vars(`SLX-ID`, `Block ID`), list(as.character)) 
})) %>% mutate(Samplename = paste(`SLX-ID`,`Index Sequence`, sep='.')) %>% 
  rowwise %>% mutate_at(vars(`SLX-ID`), list(pastefun) ) %>% ungroup %>% 
  mutate(
  `Hospital Research ID` = str_replace_all( str_remove_all(`Hospital Research ID`, " "), '/', '_'), 
  `Index Sequence` = str_replace_all(`Index Sequence`, 'tp', ''),
  Samplename = paste(`SLX-ID`,`Index Sequence`,sep='.')
  )

@@

@@

## Normalize validation cohort within itself

Instead scaling using z-values from the trianing data, I can do it from within the validation cohort. This looks better here, but when I use this for predictions it doesn't significantly change the results.

@@

```{r}
preds = do.call(bind_rows, purrr::map(
  list.files('~/Data/BarrettsProgressionRisk/Analysis/validation/multipcf/predictions_scale_within_cohort/0.9/', pattern='predictions.tsv', recursive = T, full.names = T), 
  function(f) read_tsv(f, col_types = 'cddcccccccc'))) %>% mutate(Endoscopy = as.Date(Endoscopy)) %>% dplyr::select(-`Path Notes`) %>% mutate(`Hospital Research ID` = str_replace_all( str_remove_all(`Hospital Research ID`, " "), '/', '_'))

preds = left_join(preds,all.val, by=c('Sample'='Samplename', 'Hospital Research ID', 'SLX-ID', 'Index Sequence', 'Block ID'))

# FPR is ~40% still
nrow(preds %>% filter(Status == 'NP' & Risk == 'High'))/nrow(preds %>% filter(Status == 'NP'))

@@

preds = left_join(preds,all.val, by=c('Sample'='Samplename', 'Hospital Research ID', 'SLX-ID', 'Index Sequence', 'Block ID'))

@@

How does it change if I use a global variance learned from both cohorts, or a rolling-bin variance (to preserve some local info) to adjust?

I think a weight-adjusted global~local variance isn't going to help.  The variance is consistently and stubbornly greater in the validation cohort no matter which direction I weight the scale factor.

```{r}
wt = c(0.5,0.5)

# global variance
all.tiles = bind_rows(training.tiles, val.tiles)
global.sd = sd(as.matrix(all.tiles[,-1]))

means = apply(training.tiles[,-1], 2, mean)
sds = apply(training.tiles[,-1], 2, sd)
wt.sds = sapply(sds, function(x) weighted.mean(c(x,global.sd),w=wt))

@@

means = apply(val.tiles[,-1], 2, mean)
sds = apply(val.tiles[,-1], 2, sd)
wt.sds = sapply(sds, function(x) weighted.mean(c(x,global.sd),w=wt))

@@

means = apply(all.tiles[,-1], 2, mean)
sds = apply(all.tiles[,-1], 2, sd)
wt.sds = sapply(sds, function(x) weighted.mean(c(x,global.sd),w=wt))

@@

```{r eval=F}

@@

#From training data
plot(tx.normal)

#From validation data
plot(vx.normal)
```

```{r eval=F}

### 2019-09-10 Validation Report.Rmd
...

val.file = '~/Data/BarrettsProgressionRisk/QDNAseq/validation/sWGS_validation_batches.xlsx'
analysis.dir = '~/Data/BarrettsProgressionRisk/Analysis/validation/multipcf'
preds.dir = paste0(analysis.dir, '/predictions/0.9')

>>>>

val.file = '~/Data/BarrettsProgressionRisk/QDNAseq/validation/sWGS_validation_batches.xlsx'
analysis.dir = '~/Data/BarrettsProgressionRisk/Analysis/validation/multipcf'
preds.dir = paste0(analysis.dir, '/predictions_model_5e6/0.9')
